Evolution of precopulatory and postcopulatory inbreeding behaviour

### 2016-09-07 ms_pre_post.Rmd
Revise Model section to fix errors and incorporate new changes to the model analysis structure

---
title: "Evolution of precopulatory and postcopulatory inbreeding behaviour"
author: "A. Bradley Duthie et al."
date: "August 24, 2016"
output:
  pdf_document:
    fig_caption: yes
  html_document: default
  word_document:
    fig_caption: yes
    pandoc_args:
    - --csl
    - evolution.csl
header-includes:
- \usepackage{natbib}
- \usepackage{lineno}
- \linenumbers
- \bibliographystyle{amnatnat}
linestretch: 1
csl: evolution.csl
bibliography: ms_pre_post.bib
---

@@

We track interactions among discrete individuals in a small focal population that persists over multiple generations. Individuals within this population are either female or male, and are potentially related if they share a common ancestory. We record the coefficient of kinship ($k$), defined as the probability that a randomly selected homologous allele from each individual is identical-by-descent [@Lynch1998], between all pairs of individuals in each generation using the population pedigree [see @Boyce1983; @Duthie]. At the start of each generation, each female pays up to three independent costs of (1) tendency to engage in polyandry ($c_{P}$), (2) mating strategy with respect to inbreeding ($c_{M}$), and (3) fertilisation strategy ($c_{F}$) with respect to inbreeding. The number of males that a focal female mates with ($N_{males}$) is determined by the effects of 20 alleles ($P_{a}$) across 10 diploid loci that combine additively to determine a her genotype ($P_{g}$), and ultimately the phenotype for her tendency to engage in polandry ($P_{p}$). A focal female's choice of mate(s) is affected by 20 alleles ($M_{a}$) across 10 diploid loci that combine additively to affect her genotype ($M_{g}$) and phenotype ($M_{p}$) for pre-copulatory inbreeding avoidance or preference, and $k$ between the focal female and her potential mates. Similarly, after mating has occurred, post-copulatory fertilisation of a focal female's $n$ total offspring is affected by 20 alleles ($F_{a}$) across 10 diploid loci that combine additively to affect the genotype ($F_{g}$) and phenotype ($F_{p}$) of the degree to which fertilisation success is affected by $k$. After fertilisation, the vialbility of a focal female's $n$ offspring is decreased by inbreeding depression ($\beta$); surviving offspring become adults. After inbreeding depression in offspring, $\rho$ adult male immigrants arrive in the population, then density effects reduce the entire population down to female ($K_{f}$) and male ($K_{m}$) carrying capacity. Over generations, we record mean values of alleles underlying polyandry, mating strategy, and fertilisation strategy.

@@

<!--- Change mating cost to an increased risk of mating failure ---> 
Values of phenotypic traits for tendency to engage in polyandry ($P_{p}$), mating ($M_{p}$), and fertilistation ($F_{p}$) each have costs, which independently increase the probability of a focal female's reproductive failure. These costs are all expressed prior to mating. In the case of tendency to engage in polyandry, the cost of $P_{p}$ ($c_{P}$) is interpreted biologically as a cost associated with mate searching or courtship [e.g., @Rowe1988; @Rowe1994; @Ronkainen1994; @Koga1998; @Lasley-Rasher2012]. For inbreeding mating strategy, the cost of $M_{p}$ ($c_{M}$) is interpreted biologically as a cost associated with risk of mating failure due to females being overly choosy [e.g., @Kokko2013], and for inbreeding fertilisation strategy, the cost of $F_{p}$ ($c_{F}$) is interpreted as an energetic cost associated with the physiological or biochemical mechanisms that need to be developed to bias fertilisation for inbreeding avoidance or inbreeding preference [e.g., @Gasparini2011]. Each cost independently increases the risk of female pre-mating mortality and, because generations are non-overlapping, total reproductive failure as a linear function of its associated trait. The probability that a focal female dies before mating due to the cost of polyandry is therefore $P_{p} \times c_{P}$. Similarly, the probability of pre-mating female mortality due to the cost of mating strategy and fertilisation strategy is $|M_{p}| \times c_{M}$ and $|F_{p}| \times c_{F}$, repectively, where $|M_{p}|$ is the absolute value of $M_{p}$ and $|F_{p}|$ is the absolute value of $F_{p}$. Absolute values are used for applying costs to inbreeding strategies because both negative and positive $M_{p}$ and $F_{p}$ bias breeding interactions with respect to kinship, while only positive values of $P_{p}$ affect a female's tendency for polyandry; zero and negative $P_{p}$ values always translate to monandrous phenotypes (see below).

@@

Each female chooses a number of males to mate with ($N_{males}$) based on her genetically determined tendency to engage in polyandry ($P_{p}$). The exact value of $N_{males}$ is calculated by sampling from a random poisson distribution such that $N_{males} = Poisson(P_{p}) + 1$. This random sampling procedure causes some stochastic variation in the number of realised mates a that female chooses, with a mean $N_{males}$ of $P_{p}+1$. A one is added to $P_{p}$ to ensure that all females choose at least one mate, as low values of $P_{p}$ might return $Poisson(P_{p})$ values of zero.

@@

Each male within the population is a potential mate for any female to choose. We therefore assume that there are no opportunity costs associated with male mating such that mating with one female would reduce a male's ability to mate with and sire the offspring of any other female [e.g., @Waser1986]. While all males are therefore available as potential mates, we assume that a female mates with her total allotment of $N_{males}$ without replacement, meaning that $N_{males}$ models the total number of distinct males that a female chooses to mate with rather than the number of times that she mates. In the unlikely event that a female's $N_{males}$ exceeds the total number of males in the population, then she mates with all of the males in the populaton. Most often, $N_{males}$ will be a small subset of the available males in the population [@Duthie], so a focal female will be able to preferentially mate with males that are close kin (high $k$), distant kin (low $k$), or non-kin ($k=0$) through the expression of her mating phenotype ($M_{p}$). Like polyandry allele values, mating strategy allele values ($M_{a}$) can take any real number. However, unlike polyandry, the sum of all of a female's 20 $M_{a}$ alleles equals both her mating strategy genotype ($M_{g}$) and her mating strategy phenotype ($M_{p}$). Negative or positive values of $M_{p}$ cause a female to avoid or prefer mating with kin, respectively. Values of $M_{p}=0$ causes females to mate randomly with respect to kinship.

The probability that a focal female $i$ mates with a male $j$ with which she shares some kinship $k_{ij}$ is calculated by first assigning each male a perceived mate quality $Q^{m}_{ij}$. This quality depends upon whether the female has a strategy of avoiding mating with kin ($M_{p}<0$) or preferring mating with kin ($M_{p}>0$; if $M_{p}=0$, all males are assigned a quality of 1, $Q^{m}_{ij}=1$). If a focal female avoids mating with kin, then $Q^{m}_{ij} = (-M_{p} \times k_{ij} + 1)^{-1}$, meaning that the perceived quality of a male decreases lineary with increasingly positive values of $k_{ij}$ and increasingly negative values of $M_{p}$. If a focal female prefers mating with kin, then $Q^{m}_{ij} = M_{p} \times k_{ij} + 1$, meaning that the perceived quality of a male increases with increasingly positive values of $k_{ij}$ and $M_{p}$. After a focal female assigns all males a perceived quality, the sum of all male perceived qualities is calculated, then each male quality is divided by this sum. This transformation results in each male having a relative perceived mate quality $q^{m}_{ij}$, which is constrained to values between zero and one. The value of $q^{m}_{ij}$ then defines the probability that a female chooses the male as a mate; female choice is therefore stochastic, meaning that females do not always choose the male of the highest $q^{m}_{ij}$. For polyandrous females that choose more than one male mate, mates are chosen iteratively such that $Q^{m}_{ij}$ and $q^{m}_{ij}$ are re-calculated for each additional mate choice, and with $Q^{m}_{ij}$ and therefore $q^{m}_{ij}$ values of already chosen males set to zero to ensure mate sampling without replacement.

@@

The probability that an offspring of a focal female $i$ is fertilised by one of $i$'s mates $j$ is calculated by first assigning a perceived fertilisation quality to each $j$, $Q^{f}_{ij}$. Perceived fertilisation quality $Q^{f}_{ij}$ is calculated in the same way as perceived mate quality $Q^{m}_{ij}$, such that if a focal female has a fertilisation strategy of avoiding inbreeding ($F_{p}<0$), then the perceived quality of male $j$ is $Q^{f}_{ij} = (-F_{p} \times k_{ij} + 1)^{-1}$. If instead a focal female has a fertilisation strategy of preferring inbreeding ($F_{p}>0$), then the perceived quality of male $j$ is $Q^{f}_{ij} = F_{p} \times k_{ij} + 1$. The perceived qualities of every male that a focal female has mated with are summed, then a relative quality ($q^{f}_{ij}$) is calculated for each male by dividing each individual male mate's $Q^{f}_{ij}$ by this sum. As with relative perceived mate quality values ($q^{m}_{ij}$), this results in values of perceived fertilisation quality $q^{f}_{ij}$ between zero and one. These $q^{f}_{ij}$ values define the probability that the offspring of a focal female $i$ is fertilised by the male $j$. Females produce $n$ offspring, meaning that $q^{f}_{ij}$ values are sampled $n$ times for each female to determine the realised distribution of sires.

@@

Alleles of a focal female's offspring mutate with a probability of $\mu=0.001$, with mutation being an independent event across loci. In general, loss of function mutations might be expected to be more frequent than mutations that increase the functionality of adaptations that rely on complex physiologies or behaviours. If tendency for polyandry, inbreeding mating strategies, or inbreeding fertilisation strategies rely on such complex physiologies or behaviours, then we might expect the direction of mutation to be biased (e.g., towards monandry or random mating). However, because we are interested specifically in how selection affects reproductive strategy given varying costs and inbreeding depression, we model mutation as unbiased so that selection can be isolated and inferred clearly from mean allele values.

Hence, when a mutation occurs at a loci, a new value is sampled from a random normal distribution with a mean of zero, $\mathcal{N}(0, \mu_{SD})$. This new value is then added to the original value of the allele that mutates [@Kimura1965; @Lande1976; @Bocedi2014; @Duthie]. The standard deviation of the mutation effect size ($\mu_{SD}$) is set to $\sqrt{1/20}$, which scales the variance of mutation effect size to the total number of alleles affecting $P_{p}$, $M_{p}$, and $F_{p}$ phenotypes [see @Kimura1965; @Lande1976; @Bocedi2014].

@@

The viability of a focal female $i$'s offspring ($\Psi_{\textrm{off}}$) decreases as a function of the kinship coefficient between $i$ and the sire of her offspring $j$ ($k_{ij}$), and a log-linear slope of inbreeding depression $\beta$. We model selection as having an absolute effect on offspring viability (i.e., hard selection), rather than a relative effect on viability (i.e., soft selection), so that the effect of $\beta$ is consistent over generations and across different parameter combinations. Natural selection on tendency for polyandry ($P_{p}$), inbreeding mating strategy ($M_{p}$), and inbreeding fertilisation strategy ($F_{p}$) can therefore be compared with respect to log-linear inbreeding depression slopes across generations and parameter combinations. If offspring viability was instead relative such that the least inbred offspring had a high viability regardless of the absolute value of $k_{ij}$, then the effect of kinship coefficients and inbreeding depression slopes on offspring viability would depend on the population-wide distribution of $k_{ij}$ between females and males, making model inference more difficult. Therefore, we define offspring viability as,
\begin{equation}
\Psi_{\textrm{off}} = e^{-\beta k_{ij}}
\end{equation}
The slope $\beta$ models the number of haploid lethal equivalents that exist in the gametes of $i$ and $j$, and which might combine to become homozygous in offspring and reduce viability. Equation 1 assumes that each deleterious recessive allele affects offspring viability independently, meaning that every deleterious recessive allele added will have a multiplicative affect on decreasing viability [@Morton1956; @Mills1994], causing the log-linear relationship between parent kinship and inbreeding depression. We assume that inbreeding depression always decreases offspring viability, $\beta \geq 0$ (i.e., no outbreeding depression). Therefore, because $k_{ij}$ is constrained to values between zero and one, the product of $-\beta$ and $k_{ij}$ must be a value less than or equal to zero, $-\beta \times k_{ij} \leq 0$. Values of $\Psi_{\textrm{off}}$ must therefore be between zero and one. We can therefore define $\Psi_{\textrm{off}}$ as the probability that an offspring is viable, and sample its realised viability using a Bernoulli trial.

@@

To avoid indefinite population growth over generations, we introduce a separate carrying capacity on the total number of females ($K_{f}$) and males ($K_{m}$) in the focal population following immigration [@Guillaume2009; @Duthie]. If at the end of a generation the number of females or males exceeds their respective carrying capacities, then each sex is reduced back to its $K_{f}$ or $K_{m}$ by randomly removing females and males until each sex is at its carrying capacity. Removal of females and males can be interpreted as dispersal outside the focal population, mortality, or as some combination of dispersal and mortality. The remaining females and males within the population form the next generation of adults.

@@

To analyse our model, we track mean values of alleles underlying tendency for polyandry ($P_{a}$), inbreeding mating strategy ($M_{a}$), and inbreeding fertilisation strategy ($F_{a}$) in a focal population. We present evolutionary dynamics of these mean allele values across different combinations of inbreeding depression slopes and costs to test our four focal hypotheses (see below). The combinations of costs and inbreeding depression slopes that we consider include all eight possible permutations of values of costs of tendency for polyandry ($c_{P}$), inbreeding mating strategy ($c_{M}$), and inbreeding fertilisation strategy ($c_{F}$) equal to either $0$ or $0.02$ [@Duthie observed $0.02$ to be strong cost of tendency for polyandry], and for values of $\beta = 3$ ($8$ total combinations). In all of these simulations, we record mean values of $P_{a}$, $M_{a}$, $F_{a}$ in each generation. Each of the $8$ possible cost combinations is replicated $40$ times, and grand mean values and standard deviation of means is calculated in each generation across replicate simulations. These analyses allowed us to infer how allele values change over generations in response to costs, but also in response to the changing values of other alleles and therefore potentially interacting phenotypes (e.g., selection on inbreeding fertilisation strategy might be weak if polyandry is rare).

To test our hypothesis that selection for evolution of polyandry is driven by female's ability to bias fertilisation toward inbreeding avoidance through post-copulatory mechanisms, we compare simulations in which all three phenotypes including tendency for polyandry ($P_{p}$), inbreeding mating strategy ($M_{p}$), and inbreeding fertilisation strategy ($F_{p}$) evolve, with an identical set of simulations in which $F_{p}$ cannot evolve. In simulations without evolving $F_{p}$, we therefore include all 8 combinations of $c_{P}=\{0,0.02\}$, $c_{M}=\{0,0.02\}$, $c_{F}=\{0,0.02\}$, and $\beta = 3$ replicated 40 times each. To stop $F_{p}$ from evolving, we sever the connection from $F_{a}$ to $F_{p}$ such that all $F_{g}$ genotypes resulting from $F_{a}$ cause random fertilisation with respect to kinship. Hence, while mutation and drift can still occur for $F_{a}$ alleles, these alleles have no phenotypic effect and polyandry cannot cause polyandrous females to benefit through an increased ability to bias fertilisation toward inbreeding avoidance (under such conditions, there can also never be a cost of $F_{p}$, meaning that $c_{F}=0$). By comparing evolution of mean $P_{p}$ values over generations when $F_{p}$ also evolves versus when $F_{p}=0$ (random fertilisation), we test whether or not evolution of inbreeding fertilisation strategy increases selection for evolution of polyandry.

To test hypotheses regarding evolution under different relative costs of phenotypes, we consider all combinations of costs and inbreeding depression parameter values using only simulations in which all three traits evolve. To test our hypothesis that inbreeding mating strategy will evolve toward inbreeding avoidance ($M_{p}<0$) when mating costs are low relative to the costs of tendency for polyandry and inbreeding fertilisation strategy, but that selection for inbreeding fertilisation strategy will then be negligible ($F_{p} \approx 0$), we quantify evolution of $M_{p}$ and $F_{p}$ over generations in simulations where $c_{M}=0$ but $c_{P}$, $c_{F}$, or both equals $0.02$ and $\beta=3$. We compare evolution in these three cost parameter value combinations with simulations in which all costs equal either $0$ or $0.02$. If our hypothesis is supported, then $M_{p}$ but not $F_{p}$ will evolve to more negative values when $c_{M}$ is relatively low than when all cost values are uniform. To test the hypothesis that inbreeding fertilisation strategy will evolve toward inbreeding avoidance ($F_{p}<0$) given low costs of polyandry and inbreeding fertilisation strategy relative to the cost of inbreeding mating strategy, but that selection on inbreeding mating strategy will be negligible ($M_{p} \approx 0$), we qualtify evolution of $M_{p}$ and $F_{p}$ over generations when $c_{F}=0$, $c_{P}=0$, and $c_{M}=0.02$ and $\beta=3$. We compare evolution for this combination of cost parameter values with simulations in which all costs equal either $0$ or $0.02$. If our hypothesis is supported, then $F_{p}$ but not $M_{p}$ will evolve to more negative values when $c_{F}$ is relatively low than when all cost values are uniform.

To test the hypothesis that constraints on evolution of one inbreeding strategy will increase the strength of selection on a complementary strategy, we fixed inbreeding strategy phenotypes to values of zero, thereby causing either random mating ($M_{p}=0$) or the random fertilisation of offspring ($F_{p}=0$) according to a fair raffle among mates with respect to kinship. We then compare evolution of the non-fixed trait to its evolution in simulations where all traits are allowed to evolve. In all simulations testing this hypothesis, we used strong inbreeding depression ($\beta=3$) and included $40$ replicates for each set of parameter values.

>>>>

---
title: "Evolution of precopulatory and postcopulatory inbreeding behaviour"
author: "A. Bradley Duthie et al."
date: "August 24, 2016"
output:
  word_document:
    fig_caption: yes
    pandoc_args:
    - --csl
    - evolution.csl
  pdf_document:
    fig_caption: yes
  html_document: default
header-includes:
- \usepackage{natbib}
- \usepackage{lineno}
- \linenumbers
- \bibliographystyle{amnatnat}
linestretch: 1
csl: evolution.csl
bibliography: ms_pre_post.bib
---

@@

We track interactions among discrete individuals in a small focal population that persists over multiple generations. Individuals within this population are either female or male, and can be related to each other if they share a common ancestor. We record the coefficient of kinship ($k$), defined as the probability that a randomly selected homologous allele from each individual is identical-by-descent [@Lynch1998], between all pairs of individuals in each generation using the population pedigree [see @Boyce1983; @Duthie]. At the start of each generation, each female pays up to three independent costs of (1) tendency to engage in polyandry ($c_{P}$), (2) mating strategy with respect to inbreeding ($c_{M}$), and (3) fertilisation strategy with respect to inbreeding ($c_{F}$). The number of males that a focal female mates with ($N_{males}$) is determined by the effects of 20 alleles ($P_{a}$) across 10 diploid loci that combine additively to determine her genotype ($P_{g}$), and ultimately her phenotype for tendency to engage in polandry ($P_{p}$). A focal female's choice of mate(s) is affected by 20 alleles ($M_{a}$) across 10 diploid loci that combine additively to affect her genotype ($M_{g}$) and phenotype ($M_{p}$) for pre-copulatory inbreeding avoidance or preference, and kinship between the focal female and her potential mates. Similarly, after mating has occurred, post-copulatory fertilisation of a focal female's $n$ total offspring is affected by 20 alleles ($F_{a}$) across 10 diploid loci that combine additively to affect the genotype ($F_{g}$) and phenotype ($F_{p}$) of the degree to which fertilisation success is affected by kinship. After fertilisation, the viability of a focal female's $n$ offspring is decreased by inbreeding depression ($\beta$), and surviving offspring become adults. After inbreeding depression in offspring, $\rho$ adult male immigrants arrive in the population, then density effects reduce the entire population down to female ($K_{f}$) and male ($K_{m}$) carrying capacity. Over generations, we record mean values of alleles underlying polyandry, inbreeding mating strategy, and inbreeding fertilisation strategy.

@@

<!--- Change mating cost to an increased risk of mating failure ---> 
Values of phenotypic traits for tendency to engage in polyandry ($P_{p}$), inbreeding mating strategy ($M_{p}$), and inbreeding fertilistation strategy ($F_{p}$) each have costs, which independently increase the probability of a focal female's reproductive failure. These costs are all expressed prior to mating. In the case of tendency to engage in polyandry, the cost of $P_{p}$ ($c_{P}$) is interpreted biologically as a cost associated with mate searching or courtship [e.g., @Rowe1988; @Rowe1994; @Ronkainen1994; @Koga1998; @Lasley-Rasher2012]. For inbreeding mating strategy, the cost of $M_{p}$ ($c_{M}$) is interpreted biologically as a cost associated with risk of mating failure due to females being overly choosy [e.g., @Kokko2013], and for inbreeding fertilisation strategy, the cost of $F_{p}$ ($c_{F}$) is interpreted as an energetic cost associated with the physiological or biochemical mechanisms that need to be developed to bias fertilisation for inbreeding avoidance or inbreeding preference [e.g., @Gasparini2011]. Each cost independently increases the risk of female pre-mating mortality and, because generations are non-overlapping, total reproductive failure as a linear function of its associated trait. The probability that a focal female dies before mating due to the cost of polyandry is therefore $P_{p} \times c_{P}$. Similarly, the probability of pre-mating female mortality due to the cost of inbreeding mating strategy and inbreeding fertilisation strategy is $|M_{p}| \times c_{M}$ and $|F_{p}| \times c_{F}$, repectively, where $|M_{p}|$ is the absolute value of $M_{p}$ and $|F_{p}|$ is the absolute value of $F_{p}$. Absolute values are used for applying costs to inbreeding strategies because both negative and positive $M_{p}$ and $F_{p}$ bias breeding interactions with respect to kinship, while only positive values of $P_{p}$ affect a female's tendency for polyandry; zero and negative $P_{p}$ values always translate to monandrous phenotypes (see below).

@@

Each female chooses a number of males to mate with ($N_{males}$) based on her genetically determined tendency to engage in polyandry ($P_{p}$). The exact value of $N_{males}$ is calculated by randomly sampling from a poisson distribution such that $N_{males} = Poisson(P_{p}) + 1$. This random sampling procedure causes some stochastic variation in the number of realised mates a that female chooses, with a mean $N_{males}$ of $P_{p}+1$. A value of one is added to $P_{p}$ to ensure that all females choose at least one mate, as low values of $P_{p}$ might return $Poisson(P_{p})$ values of zero.

@@

Each male within the population is a potential mate for any female to choose. We therefore assume that there are no opportunity costs associated with male mating such that mating with one female would reduce a male's ability to mate with and sire the offspring of any other female [e.g., @Waser1986]. While all males are therefore available as potential mates, we assume that a female mates with her total allotment of $N_{males}$ without replacement, meaning that $N_{males}$ models the total number of distinct males that a female chooses to mate with rather than the number of times that she mates. In the unlikely event that a female's $N_{males}$ exceeds the total number of males in the population, then she mates with all of the males in the populaton. Most often, $N_{males}$ will be a small subset of the available males in the population [@Duthie], so a focal female will be able to preferentially mate with males that are close kin (high $k$), distant kin (low $k$), or non-kin ($k=0$) through the expression of her inbreeding mating strategy phenotype ($M_{p}$). Like polyandry allele values, inbreeding mating strategy allele values ($M_{a}$) can take any real number. However, unlike polyandry, the sum of all of a female's 20 $M_{a}$ alleles equals both her inbreeding mating strategy genotype ($M_{g}$) and her inbreeding mating strategy phenotype ($M_{p}$). Negative or positive values of $M_{p}$ cause a female to avoid or prefer mating with kin, respectively. Values of $M_{p}=0$ causes females to mate randomly with respect to kinship.

The probability that a focal female $i$ mates with a male $j$ with which she shares some kinship $k_{ij}$ is calculated by first assigning each male a perceived mate quality $Q^{m}_{ij}$. This quality depends upon whether the female has a strategy of avoiding mating with kin ($M_{p}<0$) or preferring mating with kin ($M_{p}>0$; if $M_{p}=0$, all males are assigned a quality of 1, $Q^{m}_{ij}=1$). If a focal female avoids mating with kin, then $Q^{m}_{ij} = (-M_{p} \times k_{ij} + 1)^{-1}$, meaning that the perceived quality of a male decreases linearly with increasingly positive values of $k_{ij}$ and increasingly negative values of $M_{p}$. If a focal female prefers mating with kin, then $Q^{m}_{ij} = M_{p} \times k_{ij} + 1$, meaning that the perceived quality of a male increases with increasingly positive values of $k_{ij}$ and $M_{p}$. After a focal female assigns all males a perceived quality, the sum of all male perceived qualities is calculated, then each male quality is divided by this sum. This transformation results in each male having a relative perceived mate quality $q^{m}_{ij}$, which is constrained to values between zero and one. The value of $q^{m}_{ij}$ then defines the probability that a female chooses the male as a mate; female choice is therefore stochastic, meaning that females do not always choose the male of the highest $q^{m}_{ij}$. For polyandrous females that choose more than one male mate, mates are chosen iteratively such that $Q^{m}_{ij}$ and $q^{m}_{ij}$ are re-calculated for each additional mate choice, and with $Q^{m}_{ij}$ and therefore $q^{m}_{ij}$ values of already chosen males set to zero to ensure mate sampling without replacement.

@@

The probability that an offspring of a focal female $i$ is fertilised by one of $i$'s mates $j$ is calculated by first assigning a perceived fertilisation quality to each $j$, $Q^{f}_{ij}$. Perceived fertilisation quality $Q^{f}_{ij}$ is calculated in the same way as perceived mate quality $Q^{m}_{ij}$, such that if a focal female has a fertilisation strategy of avoiding inbreeding ($F_{p}<0$), then the perceived quality of male $j$ is $Q^{f}_{ij} = (-F_{p} \times k_{ij} + 1)^{-1}$. If instead a focal female has a fertilisation strategy of preferring inbreeding ($F_{p}>0$), then the perceived quality of male $j$ is $Q^{f}_{ij} = F_{p} \times k_{ij} + 1$. The perceived qualities of every male that a focal female has mated with are summed, then a relative quality ($q^{f}_{ij}$) is calculated for each male by dividing each individual male mate's $Q^{f}_{ij}$ by this sum. As with relative perceived mate quality values ($q^{m}_{ij}$), this results in values of perceived fertilisation quality $q^{f}_{ij}$ between zero and one. These $q^{f}_{ij}$ values define the probability that the offspring of a focal female $i$ is fertilised by the male $j$. Females produce $n$ offspring, meaning that $q^{f}_{ij}$ values are sampled $n$ times independently and with replacement for each female to determine the realised distribution of sires.

@@

Alleles of a focal female's offspring mutate with a probability of $\mu=0.001$, with mutation being an independent event the does or does not occur for each locus. In general, loss of function mutations might be expected to be more frequent than mutations that increase the functionality of adaptations that rely on complex physiologies or behaviours. If tendency for polyandry, inbreeding mating strategies, or inbreeding fertilisation strategies rely on such complex physiologies or behaviours, then we might expect the direction of mutation to be biased (e.g., towards monandry or random mating). However, because we are interested specifically in how selection affects reproductive strategy given varying costs and inbreeding depression, we model mutation as unbiased so that selection can be isolated and inferred clearly from mean allele values.

Hence, when a mutation occurs at a locus, a new value is sampled from a random normal distribution with a mean of zero and a standard devation of mutation effect size $\mu_{SD}$, $\mathcal{N}(0, \mu_{SD})$. This new value is then added to the original value of the allele that mutates [@Kimura1965; @Lande1976; @Bocedi2014; @Duthie]. The standard deviation of the mutation effect size ($\mu_{SD}$) is set to $\sqrt{1/20}$, which scales the variance of mutation effect size to the total number of alleles affecting $P_{p}$, $M_{p}$, and $F_{p}$ phenotypes [see @Kimura1965; @Lande1976; @Bocedi2014].

@@

The viability of a focal female $i$'s offspring ($\Psi_{\textrm{off}}$) decreases as a function of the kinship coefficient between $i$ and the sire of her offspring $j$ ($k_{ij}$), and a log-linear slope of inbreeding depression $\beta$. We model selection as having an absolute effect on offspring viability (i.e., hard selection), rather than a relative effect on viability (i.e., soft selection), so that the effect of $\beta$ is consistent over generations and across different parameter combinations. Natural selection on tendency for polyandry ($P_{p}$), inbreeding mating strategy ($M_{p}$), and inbreeding fertilisation strategy ($F_{p}$) can therefore be compared with respect to log-linear inbreeding depression slopes across generations and parameter combinations. If offspring viability was instead relative such that the least inbred offspring had a high viability regardless of the absolute value of $k_{ij}$, then the effect of kinship coefficients and inbreeding depression slopes on offspring viability would depend on the population-wide distribution of $k_{ij}$ between females and males, making model inference more difficult. Therefore, we define offspring viability as,
\begin{equation}
\Psi_{\textrm{off}} = e^{-\beta k_{ij}}
\end{equation}
The slope $\beta$ models the number of haploid lethal equivalents that exist in the gametes of $i$ and $j$, and which might combine to become homozygous in offspring and reduce viability. Equation 1 assumes that each deleterious recessive allele affects offspring viability independently, meaning that every deleterious recessive allele added will have a multiplicative affect on decreasing total viability [@Morton1956; @Mills1994], causing the log-linear relationship between parent kinship and inbreeding depression. We assume that inbreeding depression always decreases offspring viability, $\beta \geq 0$ (i.e., no outbreeding depression). Therefore, because $k_{ij}$ is constrained to values between zero and one, the product of $-\beta$ and $k_{ij}$ must be a value less than or equal to zero, $-\beta \times k_{ij} \leq 0$. Values of $\Psi_{\textrm{off}}$ must therefore be between zero (if $-\beta \times k_{ij}$ is very negative) and one (if $-\beta \times k_{ij} = 0$). We can therefore define $\Psi_{\textrm{off}}$ as the probability that an offspring is viable, and sample its realised viability using a Bernoulli trial.

@@

To avoid having unrestricted population growth over generations, we introduce a separate carrying capacity on the total number of females ($K_{f}$) and males ($K_{m}$) in the focal population following immigration [@Guillaume2009; @Duthie]. If at the end of a generation the number of females or males exceeds their respective carrying capacities, then each sex is reduced back to its $K_{f}$ or $K_{m}$ by randomly removing females and males until each sex is at its carrying capacity. Removal of females and males can be interpreted as dispersal outside the focal population, mortality, or as some combination of dispersal and mortality. The remaining females and males within the population form the next generation of adults.

@@

To analyse our model, we track mean values of alleles underlying tendency for polyandry ($P_{a}$), inbreeding mating strategy ($M_{a}$), and inbreeding fertilisation strategy ($F_{a}$) in a focal population. We present evolutionary dynamics of these mean allele values across different parameter combinations test our four focal hypotheses (see below). We simulate up to eight possible permutations of values of costs of tendency for polyandry ($c_{P}$), inbreeding mating strategy ($c_{M}$), and inbreeding fertilisation strategy ($c_{F}$) equal to either $0$ or $0.02$ [@Duthie observed $0.02$ to be a strong cost of tendency for polyandry] for values of $\beta = 3$ ($8$ total combinations). In all simulations, we record mean values of $P_{a}$, $M_{a}$, $F_{a}$ in each generation. Each of combination of parameter values simulated is replicated $40$ times, and grand mean values and standard deviation of means is calculated in each generation across replicate simulations. These analyses allowed us to infer how allele values change over generations in response to costs, but also in response to the changing values of other alleles and therefore potentially interacting phenotypes (e.g., selection on inbreeding fertilisation strategy might be weak if polyandry is rare). For all replicates, we set the maximum number of generations to 10000, which exploratory simulations and previous modelling [@Duthie] showed to be a sufficient number of generations for inferring long-term dynamics of mean allele values and therefore selection on phenotypes. For all replicates, we also set the number of immigrants per generation to $\rho=5$, which produced a range of kin and non-kin in each generation such that females could express inbreeding strategies. Females always produced $n=8$ offspring, which was sufficient to keep populations consistently near carrying capacities and avoid population extinction. Carrying capacities for females ($K_{f}$) and males ($K_{m}$) were both set to 100 because previous modelling has shown that populations of this size are small enough that mating encounters between kin occur with sufficient frequency for selection on inbreeding strategy, but populations are also not so small that selection is typically overwhelmed by genetic drift (Duthie and Reid *in press*).

To test our hypothesis that selection for evolution of polyandry is driven by female's ability to bias fertilisation toward inbreeding avoidance through post-copulatory mechanisms, we compare simulations in which all three phenotypes including tendency for polyandry ($P_{p}$), inbreeding mating strategy ($M_{p}$), and inbreeding fertilisation strategy ($F_{p}$) evolve, with an identical set of simulations in which $F_{p}$ cannot evolve. In simulations without evolving $F_{p}$, we therefore include all $4$ combinations of $c_{P}=\{0,0.02\}$ and $c_{M}=\{0,0.02\}$. To stop $F_{p}$ from evolving, we sever the connection from $F_{a}$ to $F_{p}$ such that all $F_{g}$ genotypes resulting from $F_{a}$ cause random fertilisation with respect to kinship (i.e, $F_{p}=0$). Hence, while mutation and drift can still occur for $F_{a}$ alleles, these alleles have no phenotypic effect and polyandry cannot cause polyandrous females to benefit through an increased ability to bias fertilisation toward inbreeding avoidance (under such conditions, there can also never be a cost of $F_{p}$, meaning that $c_{F}=0$). By comparing evolution of mean $P_{p}$ values over generations when $F_{p}$ also evolves versus when $F_{p}=0$ (random fertilisation), we test whether or not evolution of inbreeding fertilisation strategy increases selection for evolution of polyandry.

To test hypotheses regarding evolution under different relative costs of phenotypes, we consider all combinations of costs using simulations in which all three traits evolve and $\beta=3$. To test our hypothesis that inbreeding mating strategy will evolve toward inbreeding avoidance ($M_{p}<0$) when mating costs are low relative to the costs of tendency for polyandry and inbreeding fertilisation strategy, but that selection for inbreeding fertilisation strategy will then be negligible ($F_{p} \approx 0$), we quantify evolution of $M_{p}$ and $F_{p}$ over generations in simulations where $c_{M}=0$ but $c_{P}$, $c_{F}$, or both equals $0.02$. We compare evolution in these three cost parameter value combinations with simulations in which all costs equal either $0$ or $0.02$. If our hypothesis is supported, then $M_{p}$ but not $F_{p}$ will evolve to more negative values when $c_{M}$ is relatively low than when all cost values are uniform. To test the hypothesis that inbreeding fertilisation strategy will evolve toward inbreeding avoidance ($F_{p}<0$) given low costs of polyandry and inbreeding fertilisation strategy relative to the cost of inbreeding mating strategy, but that selection on inbreeding mating strategy will be negligible ($M_{p} \approx 0$), we quantify evolution of $M_{p}$ and $F_{p}$ over generations when $c_{F}=0$, $c_{P}=0$, and $c_{M}=0.02$ . We compare evolution for this combination of cost parameter values with simulations in which all costs equal either $0$ or $0.02$. If our hypothesis is supported, then $F_{p}$ but not $M_{p}$ will evolve to more negative values when $c_{F}$ is relatively low than when all cost values are uniform.

To test the hypothesis that constraints on evolution of one inbreeding strategy will increase the strength of selection on a complementary strategy, we fixed inbreeding strategy phenotypes to values of zero, thereby causing either random mating ($M_{p}=0$) or the random fertilisation of offspring ($F_{p}=0$) according to a fair raffle among mates with respect to kinship. We then compare evolution of the non-fixed trait to its evolution in simulations where all traits are allowed to evolve.
